---
title: "P3: Visualization of Results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
    results_glob_str: ""
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
  pdf_document:
    dev: png
    toc: true
    latex_engine: xelatex
---

Setup
=====

First, let's load some libraries we will be using and create some resuable
helper functions for parsing the output.

```{r load_libs, message=FALSE, warning=FALSE}
library('DT')
library('dplyr')
library('knitr')
library('ggplot2')
library('gplots')
library('reshape2')
library('stringr')
library('SuperLearner')
library('RColorBrewer')
library('org.Hs.eg.db')
library('GO.db')
source('R/results_helper_functions.R')

options(stringsAsFactors=FALSE,
        knitr.duplicate.label='allow',
        digits=4)
```

Results
=======

### Load SuperLearner results

```{r run1, message=FALSE}
# results_glob_str should look like '~/P3/renamed/runs/run_1/output/*.RData'
run            <- parse_p3_results(Sys.glob(results_glob_str))
coefs          <- run$coefs
cv_risks       <- run$cv_risks
var_importance <- run$var_importance
```

### Summary of SuperLearner predictor performance

```{r predictor_performance, results='asis'}
# CV risk averages across all drugs
dat <- melt(cv_risks)
colnames(dat) <- c('drug', 'learner', 'value')
dat$learner <- as.character(dat$learner)

cvrisk_averages <- dat %>% 
    group_by(learner) %>% 
    summarise(average_risk=mean(value, na.rm=TRUE))

# Coefficient averages across all drugs
dat2 <- melt(coefs)
colnames(dat2) <- c('drug', 'learner', 'value')
dat2$learner <- as.character(dat2$learner)

coef_averages <- dat2 %>% 
    group_by(learner) %>% 
    summarise(average_coef=mean(value, na.rm=TRUE)) %>%
    arrange(desc(average_coef))

combined <- merge(cvrisk_averages, coef_averages, by='learner') %>%
    arrange(desc(average_coef))

#ggplot(dat, aes(value, color=learner, fill=learner)) + 
#    geom_density(alpha=0.5)

kable(combined)
```

### Feature importance

First, let's create a vector of feature types so that we can group our 
plots this way.

```{r feature_types}
# types of predictor features
feature_types <- c('cnv_clusters', 'cnv_longest_gene', 'cnv_max_gene',
                   'cpdb_variants', 'cpdb_zscores', 'exome_variants',
                   'go_variants', 'go_zscores', 'msigdb_variants',
                   'msigdb_zscores', 'normed_counts', 'zscores_zscores')

ftype <- rep(NA, nrow(var_importance)) 
for (cls in feature_types) {
    ftype[grepl(cls, rownames(var_importance))] <- cls
}
```

#### Distribution of feature average variable importance

Median RF variable importance for each feature across all drugs.

```{r variable_importance_dist, results='asis'}
median_var_importance <- apply(var_importance, 1, median)  

# Create a feature/type mapping
median_type_importance <- as.data.frame(
    cbind(median_var_importance=as.vector(median_var_importance), type=ftype)
)

# Density grouped by feature type
ggplot(median_type_importance, aes(x=median_var_importance, 
                                   fill=type, color=type, group=type)) +
    geom_density(alpha=0.15) +
    xlab('Median Random Forest Variable Importance')

df <- median_type_importance %>% 
        group_by(type) %>%
        summarize(avg=mean(as.numeric(median_var_importance))) %>%
        arrange(desc(avg))
kable(df)
```

#### Distribution of feature maximum variable importance

```{r variable_importance_dist}
max_var_importance <- apply(var_importance, 1, max) 

# Create a feature/type mapping
max_type_importance <- as.data.frame(
    cbind(max_var_importance=as.vector(max_var_importance), type=ftype)
)

# Density grouped by feature type
ggplot(max_type_importance, aes(x=max_var_importance, 
                                 fill=type, color=type, group=type)) +
    geom_density(alpha=0.15) +
    xlab('Maximum Random Forest Variable Importance')

df <- max_type_importance %>% 
        group_by(type) %>%
        summarize(avg=mean(as.numeric(max_var_importance))) %>%
        arrange(desc(avg))
kable(df)
```

#### Features with highest average variable importance

Features ranked by _median_ random forest variable importance across all drugs.

```{r results='asis', message=FALSE}
highest_median_var_importance <- as.data.frame(sort(median_var_importance, decreasing=TRUE))
colnames(highest_median_var_importance) <- c('Importance')
highest_median_var_importance <- add_rownames(highest_median_var_importance, 'Feature')

# add human-readable annotations
highest_median_var_importance$Description = get_feature_descriptions(highest_median_var_importance$Feature)

xkable(highest_median_var_importance, 15, 
       caption='Features ranked by median variable importance',
       str_max_width=60)
```

#### Features with highest maximum variable importance

Features ranked by _max_ random forest variable importance across all drugs.

```{r results='asis', message=FALSE}
highest_max_var_importance <- as.data.frame(sort(max_var_importance, decreasing=TRUE))
colnames(highest_max_var_importance) <- c('Importance')
highest_max_var_importance <- add_rownames(highest_max_var_importance, 'Feature')

# add human-readable annotations
highest_max_var_importance$Description = get_feature_descriptions(highest_max_var_importance$Feature)

xkable(highest_max_var_importance, 15, 
       caption='Features ranked by max variable importance',
       str_max_width=60)
```

#### Features with the highest trimmed mean variable importance

Features ranked by _mean_ of their top 25 variable importance scores. Since
different features are likely to help predict responses for different drugs,
this helps us to find features which performed well for at least some subset of
the drugs.

```{r results='asis', message=FALSE}
top25 <- apply(var_importance, 1, function(x) { mean(head(sort(x, decreasing=TRUE), 25)) }) 
trimmed_mean_import <- as.data.frame(sort(top25, decreasing=TRUE))
colnames(trimmed_mean_import) <- c('Importance')
trimmed_mean_import <- add_rownames(trimmed_mean_import, 'Feature')

trimmed_mean_import$Description = get_feature_descriptions(trimmed_mean_import$Feature)

xkable(trimmed_mean_import, 15, 
       caption='Features ranked by trimmed-mean variable importance.',
       str_max_width=60)
```

### Drugs with shared predictive features

Next, we will see if there are any groups of drugs which are predicted by
similar sets of features. To do this, we will cluster the drugs based on the
their feature importance vectors.

```{r drug_heatmap, fig.width=1080/96, fig.height=1080/96, dpi=96}
heatmap_colors <- colorRampPalette(brewer.pal(9, "YlGnBu"))(100)
dist_matrix <- cor(var_importance, method='spearman')

dendrogram <- ifelse(nrow(dist_matrix) > 100, 'none', 'column')

heatmap.2(dist_matrix, trace="none", labRow=NA, dendrogram=dendrogram,
		  main="Drug prediction similarity", margins=c(12,8),
		  xlab='Drug', ylab='Drug', col=heatmap_colors)
```

