---
title: "P3: Visualization of Results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
    results_glob_str: ""
    output_dir: ""
output:
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    number_sections: true
  pdf_document:
    dev: png
    toc: true
    latex_engine: xelatex
---

Setup
-----

```{r load_libs, message=FALSE, warning=FALSE, include=FALSE}
library('DT')
library('dplyr')
library('knitr')
library('ggplot2')
library('gplots')
library('heatmaply')
library('reshape2')
library('stringr')
library('tibble')
library('SuperLearner')
library('RColorBrewer')
library('org.Hs.eg.db')
library('GO.db')
source('R/results_helper_functions.R')

options(stringsAsFactors=FALSE,
        knitr.duplicate.label='allow',
        digits=4)

# If rmarkdown.pandoc.to not specified (for example, when kniting
# piece-by-piece in Vim-R), have it default to 'latex' output.
if (is.null(opts_knit$get("rmarkdown.pandoc.to"))) {
    opts_knit$set(rmarkdown.pandoc.to='latex')
}

# For images containing unicode text, we will use cairo for PDF output
if (opts_knit$get("rmarkdown.pandoc.to") == 'latex') {
    opts_chunk$set(dev='cairo_pdf',
                   dev.args=list(cairo_pdf=list(family="DejaVu Sans")),
                   fig.width=1920/192,
                   fig.height=1920/192,
                   dpi=192)
} else {
    # HTML output
    opts_chunk$set(dev='png',
                   fig.width=1600/192,
                   fig.height=1600/192,
                   dpi=192, fig.retina=1)
}

# fix namespace for dplyr functions
assign('select', dplyr::select, envir=.GlobalEnv)
```

```{r message=FALSE}
run            <- parse_p3_results(Sys.glob(params$results_glob_str))
coefs          <- run$coefs
cv_risks       <- run$cv_risks
feature_importance <- run$feature_importance
```

Results
-------

### SuperLearner algorithm performance

#### Absolute CV risk

```{r predictor_performance, results='asis'}
# CV risk averages across all drugs
cv_risks_long <- melt(cv_risks)
colnames(cv_risks_long) <- c('drug', 'learner', 'value')
cv_risks_long$learner <- as.character(cv_risks_long$learner)

cvrisk_averages <- cv_risks_long %>%
    group_by(learner) %>%
    summarise(average_risk=mean(value, na.rm=TRUE))

# Coefficient averages across all drugs
coefs_long <- melt(coefs)
colnames(coefs_long) <- c('drug', 'learner', 'value')
coefs_long$learner <- as.character(coefs_long$learner)

coef_averages <- coefs_long %>%
    group_by(learner) %>%
    summarise(average_coef=mean(value, na.rm=TRUE)) %>%
    arrange(desc(average_coef))

# For each algorithm, how many compounds have nonzero coeficients?
coef_nonzero <- coefs_long %>% 
    group_by(learner) %>% 
    summarise(ratio_nonzero=sum(value > 0) / length(value)) %>%
    arrange(desc(ratio_nonzero))

combined <- merge(cvrisk_averages, 
                  merge(coef_averages, coef_nonzero, by='learner'),
                  by='learner') %>%
    arrange(desc(ratio_nonzero))
xkable(combined)
```

#### Relative risk (~R^2)

##### R^2 for each Learner

```{r, results='asis', fig.cap='Distribution of R^2 values', fig.width=1920/192, fig.height=1920/192}
# compute coefficient of determination (R^2) for each prediction
rsquared <- 1 - (cv_risks/cv_risks[,'mean'])
rsquared <- rsquared[,colnames(rsquared) != 'mean']

# gbmOOB_* returned NA on rare occassions
summary_df <- as.data.frame(cbind(min=apply(rsquared, 2, min, na.rm=TRUE),
                                  median=apply(rsquared, 2, median, na.rm=TRUE),
                                  max=apply(rsquared, 2, max, na.rm=TRUE))) %>%
                arrange(desc(max))
rownames(summary_df) <- colnames(rsquared)
xkable(summary_df, caption='R^2 risk by algorithm')

rsquared_long <- melt(rsquared)
colnames(rsquared_long) <- c('drug', 'learner', 'value')

# Excluding outlier gbmOOB for now (high peak at 0 messes up y limits)
ggplot(rsquared_long %>% filter(learner != 'gbmOOB'), aes(x=value, fill=learner)) + 
    geom_density(alpha=0.25) +
    geom_vline(xintercept=0, colour="red", linetype = "longdash") +
    theme(legend.justification=c(1,1), legend.position=c(1,1),
          legend.background=element_rect(fill=alpha('#333333', 0.4)),
          legend.text=element_text(size=10)) +
    xlim(-0.5, 0.5)
```

##### Number of R^2 values > 0.15 for each learner

```{r, results='asis'}
interesting_hits <- as.data.frame(apply(rsquared, 2, function (x) { sum(x > 0.15, na.rm=TRUE) }))
colnames(interesting_hits) <- 'num'
xkable(interesting_hits)
```

##### Number of NA's in CV risk output

```{r}
# learners
has_nas <- apply(cv_risks, 2, function(x) { sum(is.na(x)) })
has_nas[has_nas > 0]

# drugs
rownames(cv_risks)[apply(cv_risks, 1, function(x) { sum(is.na(x)) }) > 0]
```

#### Absolute Risk vs. Relative Risk (R^2)

```{r, warning=FALSE}
abs_vs_rel <- cbind(cv_risks_long  %>% filter(learner != 'mean') %>%
                    dplyr::rename(absolute=value),
                    relative=rsquared_long$value)

# Plot 4 at a time using facet grid
learners <- unique(abs_vs_rel$learner)
learner_groups <- split(learners, ceiling(seq_along(learners) / 4))

for (learners in learner_groups) {
    # ylimit set to avoid having a small number of extreme outliers from
    # having a drastic effect on the dynamic range of the plots.
    plt <- ggplot(abs_vs_rel %>% filter(learner %in% learners),
                  aes(x=relative, y=absolute, color=learner)) +
            geom_point(size=0.5) +
            geom_vline(xintercept=0, colour="#999999", linetype="longdash", size=0.3) +
            xlim(-0.5, 0.5) +
            ylim(0, 7.5) +
            facet_wrap(~learner, ncol=2) +
            theme(legend.position="none")
    print(plt)
}
```

### Random forest feature importance

First, let's create a vector of feature types so that we can group our
plots this way.

```{r feature_types}
# types of predictor features
feature_types <- c('cnv_clusters', 'cnv_longest_gene', 'cnv_max_gene',
                   'cpdb_variants', 'cpdb_zscores', 'exome_variants',
                   'go_variants', 'go_zscores', 'msigdb_variants',
                   'msigdb_zscores', 'normed_counts', 'zscores_zscores')

ftype <- rep(NA, nrow(feature_importance))
for (cls in feature_types) {
    ftype[grepl(cls, rownames(feature_importance))] <- cls
}
```

First, let's exclude all drugs for which none of the predictions were accurate.

```{r}
# trimmed mean function
trimmed_mean <- function(x, n=10) {
    mean(head(sort(x, decreasing=TRUE), n))
}

# sort drugs by there max and trimmed mean R^2 scores
drug_scores <- data_frame(drug=rownames(rsquared), 
                          max_rsquared=apply(rsquared, 1, max, na.rm=TRUE),
                          trimmed_mean_rsquared=apply(rsquared, 1, trimmed_mean)) %>%
    arrange(desc(max_rsquared + trimmed_mean_rsquared))

# exclude low-performing drugs;
# to pass the filter, drug should have either or both a max R^2 >= 0.25, or
# a trimmed mean (mean of top 10 R^2 values) >= 0.15.
top_drug_hits <- (drug_scores %>% filter(max_rsquared >= 0.25 | 
                                        trimmed_mean_rsquared >= 0.15))$drug

feature_importance_filtered <- feature_importance[,colnames(feature_importance) %in%
                                                   top_drug_hits]
```

#### Trimmed mean variable importance

Trimmed mean random forest (RF) variable importance for each feature across the top
10 drugs with an R^2 value of at least 0.25 for one or more of the learners.

##### Distribution of mean variable importance scores

Features ranked by _mean_ of their top N=10 variable importance scores. Since
different features are likely to help predict responses for different drugs,
this helps us to find features which performed well for at least some subset of
the drugs.

```{r results='asis', message=FALSE}
# Number of highest scoring drugs to average across
n <- 10

# get mean variable importance for top N compounds
top_n <- apply(feature_importance_filtered, 1, function(x) { mean(head(sort(x, decreasing=TRUE), n)) }) 

# convert to a data frame and add feature type column
trimmed_mean_import <- cbind(as.data.frame(top_n), type=ftype)
colnames(trimmed_mean_import) <- c('Importance', 'type')

trimmed_mean_import <- rownames_to_column(trimmed_mean_import, 'Feature') %>%
    arrange(desc(Importance))

trimmed_mean_import$Description = get_feature_descriptions(trimmed_mean_import$Feature)
```

```{r variable_importance_dist, results='asis'}
# Density grouped by feature type
mean_type_importance <- trimmed_mean_import %>% select(Importance, type)

ggplot(mean_type_importance, aes(x=Importance, fill=type, color=type, group=type)) +
    geom_density(alpha=0.15) +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
          legend.background=element_rect(fill=alpha('#ffffff', 0)),
          legend.justification=c(1,1), legend.position=c(1,1)) +
    xlab('Trimmed Mean Random Forest Variable Importance') +
    guides(fill=guide_legend(ncol=2))

df <- mean_type_importance %>% 
        group_by(type) %>%
        summarize(avg=mean(as.numeric(Importance))) %>%
        arrange(desc(avg))
xkable(df)
```

##### Features with the highest trimmed mean variable importance

```{r results='asis'}
xkable(trimmed_mean_import, 15, 
       caption='Features ranked by trimmed-mean variable importance.',
       str_max_width=60)

write.table(trimmed_mean_import, sep='\t', quote=FALSE,
            file=file.path(output_dir, 'trimmed_mean_featurtrimmed_mean_featuree_importance.tab'),
            row.names=FALSE)
```

#### Maximum variable importance

##### Distribution of maximum variable importance scores

```{r variable_importance_dist}
max_feature_importance <- apply(feature_importance_filtered, 1, max) 

# Create a feature/type mapping
max_type_importance <- as.data.frame(
    cbind(max_feature_importance=as.vector(max_feature_importance), type=ftype)
)

# Density grouped by feature type
ggplot(max_type_importance, aes(x=max_feature_importance,
                                 fill=type, color=type, group=type)) +
    geom_density(alpha=0.15) +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank(),
          legend.background=element_rect(fill=alpha('#ffffff', 0)),
          legend.justification=c(1,1), legend.position=c(1,1)) +
    xlab('Maximum Random Forest Variable Importance') + 
    guides(fill=guide_legend(ncol=2))

df <- max_type_importance %>%
        group_by(type) %>%
        summarize(avg=mean(as.numeric(max_feature_importance))) %>%
        arrange(desc(avg))
xkable(df)
```

##### Features with highest maximum variable importance

Features ranked by _max_ random forest variable importance across all drugs.

```{r results='asis', message=FALSE}
highest_max_feature_importance <- as.data.frame(sort(max_feature_importance, decreasing=TRUE))
colnames(highest_max_feature_importance) <- c('Importance')
highest_max_feature_importance <- rownames_to_column(highest_max_feature_importance, 'Feature')

# add human-readable annotations
highest_max_feature_importance$Description = get_feature_descriptions(highest_max_feature_importance$Feature)

xkable(highest_max_feature_importance, 15, 
       caption='Features ranked by max variable importance',
       str_max_width=60)

write.table(highest_max_feature_importance, sep='\t', quote=FALSE,
            file=file.path(output_dir, 'highest_max_feature_importance.tab'),
            row.names=FALSE)
```

### Drug response prediction performance

#### Compounds with R^2 >= 0.25

First, let's look at a list of the drugs/compounds for which P3 was able to
predict the response with a R^2 >= 0.25.

```{r}
# convert to dataframe and add max risk column
rsquared <- data.frame(rsquared)
rsquared$max_rsquared <- apply(rsquared, 1, max)

# get a table of the compounds with a R^2 of at least 0.25 for one
# or more learners 
rsquared %>% 
    rownames_to_column(var='compound') %>%
    filter(max_rsquared >= 0.25) %>%
    arrange(desc(max_rsquared)) %>%
    xkable
```

#### Top 5 drug response predictions

```{r, results='asis'}
for (i in seq_along(top_drug_hits)) {
    drug <- top_drug_hits[i]
    cat(sprintf("\n##### %d) %s\n\n", i, drug))

    # top algorithms for compound
    cat("###### Top Learners\n\n")

    df <- as.data.frame(t(rsquared[drug,]))
    colnames(df) <- 'rsquared'

    df %>% 
        rownames_to_column('learner') %>%
        arrange(desc(rsquared)) %>%
        head(15) %>%
        kable %>%
        print

    # top features (according to RF) for the compound
    cat("\n\n###### Top features\n\n")

    df <- as.data.frame(feature_importance_filtered[,drug])
    colnames(df) <- 'feature_importance'

    df %>% 
        rownames_to_column('feature') %>%
        arrange(desc(feature_importance)) %>%
        head(15) %>%
        mutate(description=get_feature_descriptions(feature)) %>%
        kable %>%
        print
}
```

#### Drugs with shared predictive features

Next, we will see if there are any groups of drugs which are predicted by
similar sets of features. To do this, we will cluster the drugs based on the
their feature importance vectors.

```{r drug_heatmap, fig.width=1080/96, fig.height=1080/96, dpi=96}
heatmap_colors <- colorRampPalette(brewer.pal(9, "YlGnBu"))(100)
dist_matrix <- cor(feature_importance_filtered, method='spearman')

dendrogram <- ifelse(nrow(dist_matrix) > 100, 'none', 'column')

heatmap.2(dist_matrix, trace="none", labRow=NA, dendrogram=dendrogram,
          main="Drug prediction similarity", margins=c(12,8),
          xlab='Drug', ylab='Drug', col=heatmap_colors)
```

### Compound vs. SuperLearner Algorithm Performance

```{r, dpi=120}
# created a filtered version of the R^2 scores for visualization
dat <- rsquared %>% 
    filter(max_rsquared >= 0.20) %>% 
    select(-max_rsquared)

# clip at [0,]
dat[dat < 0] <- 0

heatmaply(dat) %>% layout(margin=list(l=100, b=250))
```


